Visually pleasing knot projections

### 2021-09-16 inst/jma_rebuttal.tex

P10, line 17: ``... a vertical line of symmetry." - Better to say that
this figure has $D_5$ symmetry with five mirror lines.

\textcolor{blue}{Done.}

P11, line 26: It seems that all the functionality needed to deal with
links is already present in the software packages mentioned, as long
as one can let the program go around more than one single closed loop.

\textcolor{blue}{I have spent many many hours pondering how to deal
  with links.  As the referee says, much of the required functionality
  is (in principle) already present in the software.  Each component
  would have its own badness, and in addition there would be
  $n(n-1)/2$ inter-component interaction terms measuring features such
  as closeness between non-intersecting strands of different
  components.  Several problems stood out: firstly, the problem of
  inter-component points.  Currently this is a (symmetrical) Boolean
  matrix of strands with rows and columns corresponding to strands,
  and entries corresponding to whether that strand pair intersects or
  not.  The link generalization would be a symmetric matrix of
  intersection matrices: an object with four indices and entry
  (i,j,k,l) indicating whether strand j of component i intersects
  strand k of component l.  I found this object to be unweildy to work
  with and difficult to manipulate.  The second problem was the
  imposition of symmetry.  Sometimes one might desire that some
  components of a link have a particular kind of symmetry and others
  to have a different kind (or no!) symmetry.  One unexpectedly
  difficult problem was that one might have a subset of links that
  individually possess no symmetry but collectively possess mirror-,
  rotational-, or indeed dihedral- symmetry.  Thistlethwaite's L4a1 is
  problematic: one would expect the two components to be identical
  except for a 90-degree rotation, and implementing this in the
  context of the package I found impossible. L6a3 is similarly
  difficult.  We would need new badnesses too.  Consider L6a4; the
  smaller component should be not only as circular as possible (?),
  but also one might desire the overall length to be smaller than that
  of the other component.\\ My overall view was that dealing with
  links would be possible in principle but, due to a number of rather
  poorly delineated reasons (which are not really suitable for
  inclusion in a scholarly publication), an order of magnitude harder
  than the single-component links considered in the submission.  I'm
  not saying it's impossible, but just very difficult and, for me,
  firmly in the category of ``further work''.  I hope that's OK.}

### 2021-09-24 inst/jma_rebuttal.tex

The paper also needs to do a far stronger job of placing itself within
the relevant literature. Simply showing that a single diagram from
wikipedia has weaknesses is not sufficient. The nineteenth century
series of papers ``On Knots" by Peter Guthrie Tait in the Transactions
of the Royal Society of Edinburgh contain far superior images for
example. This is not the first paper to look at making pleasing knot
diagrams, the software knotplot for example offers many tools to do a
similar optimisation to the one discussed here, but is not cited,
Laura Taalman and Colin Abrams among others have multiple papers
addressing the creation of different knot representations and
optimising their form. More broadly the idea of using optimisation in
mathematical art has been used widely, the work of Robert Bosch and
Craig Kaplan springing immediately to mind.

For these reasons the current paper is not really strong enough for
publication. In order to get strong enough I would propose three major
pieces of work:

1) Conduct a thorough literature review of work on the creation of
knot diagrams and their aesthetics and place this paper into that
context as well as the broader context of optimisation in art. % issue~#39

\textcolor{blue}{I have added a literature review as suggested that
  discusses knot diagrams from early Celtic artwork through to
  mathematica.\\ \\ The referee makes the perfectly reasonable
  suggestion of discussing the broader context of optimization of art.
  I will address this comment colloquially.  It is obvious to me, and
  clearly to the referee as well, that numerical optimization
  techniques are well-suited to ``rounding off rough edges'' from a
  wide range of applications; and further that numerical optimization
  will produce ``nice looking'' results.  Is is also reasonable to
  assume that such techniques will have been used in the past by
  graphical artists.  But I can find no evidence of such activity,
  after an extensive online search and consulting with two librarians
  (mathematics and art subject specialists).  Perhaps we are using the
  wrong search terms?  I would gladly take suggestions on previous
  work to cite.  I have added a sentence to the effect that this seems
  to be new to the ``further work'' section.}

### 2021-09-24 inst/jma_knot.tex
fixes #33
fixes #39

A {\em mathematical knot} is a smooth, unoriented embedding of a
circle~$\mathbb{S}^1$ into~$\mathbb{R}^3$~\cite{manturov2004}.  Two
knots are said to be equivalent if the embedding of one can be
continuously deformed to the other; if so, there is a
homeomorphism~$h\colon\mathbb{R}^3\longrightarrow\mathbb{R}^3$ which
takes one embedded circle to the other.  In the art world, knot theory
has applications to the study of dance~\cite{khorami2020}, sculpture,
\cite{bosch2010, widmark2020}, and many other branches of mathematical
art~\cite{hart2008}.

It is common to present a knot using diagrams such as
Figure~\ref{knot_table}, in which a two-dimensional projection of the
knot is given with broken lines indicating where one strand passes
over another.  Knot diagrams figure prominently in Celtic
artwork~\cite{cromwell2008} and are discussed from a mathematical
perspective by Browne~\cite{browne2005,browne2006}.

@@

The {\tt knotR} package allows the user to create rough diagrams of
knots using the inkscape suite of software, and subsequently polish up
such diagrams in terms of a customizable objective function using
numerical optimization techniques.  Further work might include
functionality to deal with links.

@@

\begin{figure}[htbp]
  \begin{center}
\includegraphics[width=13cm]{knots_to_8crossings}  % created by inst/knotter.R
\caption{All \label{all8} prime knots with 8 or fewer crossings,
notation following Rolfsen~\cite{rolfsen2003}}
  \end{center}
\end{figure}

>>>>

A {\em mathematical knot} is a smooth, unoriented embedding of a
circle~$\mathbb{S}^1$ into~$\mathbb{R}^3$~\cite{manturov2004,adams2004}.  Two
knots are said to be equivalent if the embedding of one can be
continuously deformed to the other; if so, there is a
homeomorphism~$h\colon\mathbb{R}^3\longrightarrow\mathbb{R}^3$ which
takes one embedded circle to the other.  In the art world, knot theory
has applications to the study of dance~\cite{khorami2020}, sculpture,
\cite{bosch2010, widmark2020}, and many other branches of mathematical
art~\cite{hart2008}.

It is common to present a knot using diagrams such as
Figure~\ref{knot_table}, in which a two-dimensional projection of the
knot is given with broken lines indicating where one strand passes
over another.  Knot diagrams have a long history:
Przytycki~\cite{przytycki1998} discusses an intricate braid used as a
seal sign in around 1700\,BCE; they figure prominently in {\em The
  Book of Kells} and the Lindisfarne gospels, graphic art produced
around 800\,CE~\cite{bain1973}.  Such artwork makes extensive use of
symmetry~\cite{cromwell2008}.

Knot diagrams are ubiquitous in the modern discipline of mathematical
knot theory.  In the nineteeth century we see Tait~\cite{tait1884}
publishing an early systematic study of knots including a beautiful
plate of elegantly hand-drawn cursive knot diagrams.  The tradition
continues into the twentieth century with Rolfsen~\cite{rolfsen1976}
presenting an appendix containing many knot and link projections.
Rolfsen's diagrams are used widely today: Adams~\cite{adams2004}
reproduces his work extensively, as do many more recent works.  Online
resources are becoming more widely available, with one prominent
example being the {\em Knot Atlas}~\cite{knot_atlas} presenting
extensive knot tables following Rolfsen~\cite{rolfsen1976}---for knots
up to~10 crossings---and Thistlethwaite~\cite{hoste1998}, for knots
of~11 crossings.  Symbolic software such as Mathematica and Maple can
be used to produce representations of knots, typically in the form of
nonstandard embeddings of the torus into~$\mathcal{R}^3$; but as
discussed below this approach is suboptimal from the perspective of producing a
visually appealing projection.

@@

The {\tt knotR} package allows the user to create rough diagrams of
knots using the inkscape suite of software, and subsequently polish up
such diagrams in terms of a customizable objective function using
numerical optimization techniques.  Further work might include
functionality to deal with links.  In the broader context of
optimisation in art, we observe that numerical optimisation techniques
can produce aesthetically pleasing results, an observation that might
find uptake by graphic artists.

@@

\begin{figure}[htbp]
  \begin{center}
\includegraphics[width=13cm]{knots_to_8crossings}  % created by inst/knotter.R
\caption{All \label{all8} prime knots with 8 or fewer crossings,
notation following Rolfsen~\cite{rolfsen1976}}
  \end{center}
\end{figure}
